---
title: "scrap"
output: html_document
date: "2024-01-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r pop-agg-prep-2, warning = FALSE, eval = FALSE}

# Intersect tracts with Voronoi polygons 

<<<<<<< HEAD
# vorclip_tracts <- st_union(fn_trct, vor_chi)%>%
#   st_collection_extract("POLYGON")
# 
# vorclip_joined <- st_join(vor_chi, vorclip_tracts)
=======
vorclip_tracts <- st_union(fn_trct, vor_chi)%>%
  st_collection_extract("POLYGON")

plot(vorclip_tracts$geometry)

vorclip_joined <- st_join(vor_chi, vorclip_tracts)
>>>>>>> f18e9e24282712f425b159ebb91de62d97ef0049

# Join block centroids to new intersection polygon to get population data into the tracts again

# int_cent <- st_join(vorclip_tracts, centroids)#%>%
  #dplyr::select(TRACTCE, POP20)

# Summarize population by intersected tracts 

# int_tract <- aggregate(int_cent$POP20, by = list(int_cent$TRACTCE), FUN= sum)
# 
# int_tract$TRACTCE <- c(group_tract$Group.1)
# int_tract$SUMPOP2 <- c(group_tract$x)
# 
# int_tract2 <- int_tract %>% 
#                 dplyr::select(TRACTCE,SUMPOP2)
# 
# vor_tracts <- dplyr::left_join(vorclip_tracts, int_tract2)
# 
# plot(vor_tracts$geometry)

```

```{r map-vor-tracts}
# Map!
# qpal3 <- colorQuantile(c("#F5FDF1", "#0D5B1E"), int_grp_tract$sum_pop2, n = 10)
# 
# int_grp_tract_m <- leaflet() %>% 
#            #   addProviderTiles(providers$Esri.WorldGrayCanvas) %>% 
#            #   addPolygons(data = int_grp_tract, stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,color = ~qpal3(sum_pop2))
#               
# 
# int_grp_tract_m
```

```{r pop-agg-calc, eval = FALSE}

# Population Weighted Re-Aggregation 
 
# Calculate population weight ratio (PWR) | fragment population / source population = population weight ratio

# output <- mutate(input_data,propPop = secPOP20 / POP20)

# Multiply source fields by the PWR to estimate the population weighted value  

# output <- mutate(final_table,
          # hhi1_w = hhi1 * pwr,
          # hhi2_w = hhi2 * pwr,
          # hhi3_w = hhi3 * pwr,
          # hhi4_w = hhi4 * pwr,
          # hhi5_w = hhi5 * pwr,
          # hhi6_w = hhi6 * pwr,
          # hhi7_w = hhi7 * pwr)

```

#Scrap from 2/22/2024

## Overlays and Join to Geometry 

```{r tracts-clip-chi}
# # Use intersect overlay tool to clip all tracts in Cook County by Chicago's boundaries
# 
# chi_tracts <- st_intersection(cook_tracts, chi) %>%
#    st_collection_extract("POLYGON")%>%
#    left_join(acs_table, by= "TRACTCE")
# 
# class(chi_tracts)
# 
# plot(chi_tracts$geometry.x)

```
### Join Tract Level Data to Blocks  

My intentions for this part of the workflow were to generate a layer of centroids for the block data so I could join the population data to the tracts. I would do this two times. Once for the original tracts layer and a second time after I intersect the tracts with the voronoi diagram. I successfully generate a list of centroids for every block but when I join the centroids to the tracts layer, R makes a row for every centroid that was joined to a tract. This more than double the amount of observation in the data table. 

```{r centroids, warning = FALSE}

# Generate centroids for the blocks

cent_gen <- st_centroid(chi_pop)

centroids <- cent_gen %>%
  dplyr::select(TRACTCE20, BLOCKCE20, POP20)

# Map 
tmap_mode("plot")
tm_shape(centroids) +
    tm_dots(size=0.05, col = "black")

```

```{r centroids-to-tracts, warning = FALSE}

# Join block centroids to tracts to calculate population data for the tract

tr_cent <- st_join(chi_tracts, centroids)%>%
  dplyr::select(TRACTCE, POP20)%>%
   group_by(TRACTCE)%>%
   summarize(sumPOP = sum(POP20))


#join pop data to acs table

acs_centroids <-
 st_join(tr_cent, chi_tracts, by = "TRACTCE")%>%
  dplyr::filter(sumPOP > 0)

acs_tracts <- st_join( chi_tracts, tr_cent, by = "TRACTCE")%>%
  dplyr::filter(sumPOP > 0)

tmap_mode("plot")
tm_shape(acs_tracts) +
  tm_polygons(col="sumPOP", title="Population Tract", lwd= NA, border.col="lightgrey") +
tm_layout(legend.width=3, legend.text.size = .6, legend.title.size = .8, asp=0.8)

```

```{r lib-id}

#Join library data to centroids and select for variables that I want to keep 

lib_ids <- st_join(centroids, vor_chi)%>%
              dplyr::select(TRACTCE20, BLOCKCE20, POP20, NAME)%>%
  st_drop_geometry()

```

```{r join-trct-to-blcks}
# Join soc variables to blocks with libids


f_blocks <- left_join(lib_ids, acs_centroids, by= c("TRACTCE20" = "TRACTCE.x"))%>%
  st_as_sf()

f_blocks <- f_blocks %>%
  mutate(prPOP = POP20/sumPOP,
                    pr_minc = B19013_001E * prPOP,
                    pr_black = black * prPOP,
                    pr_for1 = for1 * prPOP) %>%
   group_by(NAME.x) %>%
   summarize(sum_black = sum(pr_black, na.rm = TRUE)),
            %>%
            st_drop_geometry()


# after I group by and summarize i just need to join that data by a unique id back to the catchment area geometry so I can map with it

s_areas <- left_join(vor_chi, f_blocks, by= c("NAME" = "NAME.x"))%>%
            dplyr::select(NAME, Census.Tracts, sum_minc, sum_black, sum_for)
              
```


```{r map-s_areas}
tmap_mode("plot")
tm_shape(s_areas) +
  tm_polygons(col="sum_minc",style= "jenks", title="Med. Household Income", lwd= 0, border.col="lightgrey") +
tm_layout(legend.width=3, legend.text.size = .6, legend.title.size = .8, asp=0.8)

tmap_mode("plot")
tm_shape(s_areas) +
  tm_polygons(col="sum_black",style= "jenks", title="Black", lwd= 0, border.col="lightgrey") +
tm_layout(legend.width=3, legend.text.size = .6, legend.title.size = .8, asp=0.8)

tmap_mode("plot")
tm_shape(s_areas) +
  tm_polygons(col="sum_for",style= "jenks", title="Foreign", lwd= 0, border.col="lightgrey") +
tm_layout(legend.width=3, legend.text.size = .6, legend.title.size = .8, asp=0.8)
```